---
title: "Assignment_2"
output: pdf_document
bibliography: 
- A2.bib
- A21.bib
- A22.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generational Shifts  
 
Traditionally, an individual’s age is one of the most common predictors of differences in attitudes and behaviors, particularly when it comes to voting. Age denotes what cohort a particular person belongs to and what point of the life cycle a particular person is in: childhood, young adult, middle aged, or retiree. Generations typically refer to people born over a 15-20-year span. Millennials are the generation born from 1980–1994 (according to the Pew Research Center), making them approximately 19-35 years old as of 2016.
 
In the United States, for approximately the past 25 years, the Baby Boomer generation, 1946–1964 was the defining vote in presidential and down ballot elections. However, according to new Pew Research Center analysis, the Millennials now outnumber the Baby Boomers, there are approximately 75.4 million Millennials, surpassing the previous generation. [1] [2] In 2016 election essentially every Millennial will be eligible to vote for the first time. 
 
## Why does this matter?
 
Millennials will play an increasingly significant role in U.S. elections as they represent an ever-greater share of eligible voters, but there are substantial differences in disposition and demographics between them and previous generations, which could well mean differences in factors that bring them to the polls. 
 
Generational shifts in attitude have been well documented.[4] Millennials are in no rush to get married, are more supportive of larger governments, are more likely to have debt, are more accepting of gay-marriage, and are more aware of gender discrimination. Additionally, most millennials came of age during a recession (@o2015millennials) and identify with and engage with technology in a way that was inconceivable to past generations. These societal shifts and cultural norms will affect how they view and partake in politics. Millennials are the only generation where just as many people describe their views as liberal as they do conservative. They also have the highest number of people reported as being open to a third party candidate and are relatively unattached to organized politics and religion.[6] 

Lastly, Millennials are diverse. First, the age range in and of itself implies a diversity of experience: an 18-year-old and a 34-year-old are, hopefully, at two very different stages in life. But more importantly, according to the Census Bureau, for every 100 millennials, 44 are a part of a minority race or ethnic group.  To put it another way, in 2000 only 8 states and Washington DC had a share of minority voters that was 30% or higher. In 2012, the number of states was 17 (plus Washington DC).[3]

## Who Votes?

The Downsian Model of Electoral competition broadly states that if policy space is unidimensional, and voters have single-peaked preferences over the policy dimension, then the median voter becomes pivotal in majority decision making. Downs (1957) used this result to analyze electoral competition, and demonstrated that two competing parties will converge to the policy of the preferred median voter. There are quite restrictive assumptions with this model, including that politicians are able to pre-commit to their proposed platforms. That being said, the model delivers a clear prediction about the impact of electoral competition on public policy and has become a mainstay in the field. 

The original Downsian Model did not pay much attention to who the "median voter"" was, so subsequent studies built on his model by showing that the median voter and median citizen are usually not one in the same. An abundant number of theoretical models have emerged, trying to explain voter turnout yet there is limited academic consensus on a core model. Different models have found empirical support in different contexts, suggesting that perhaps voter turnout has different causal mechanisms depending upon the country, voters, and time period. (Gallego, 2010; Arceneaux and Nickerson, 2009)

When it comes to the United States, non-voters have not been randomly distributed across the country. Empirical research shows that voters and nonvoters systematically differ in their socio-economic status and demographic backgrounds. These differences allude to potential differences in needs and policy preferences. (Verba, 1993; Wolfinger and Rosenstone, 1980).

A highly regarded study by Raymond Wolfinger and Steven Rosenstone (1980) demonstrated that in the US, turnout is fairly predictable based on a number of individual demographic variables. The variables are education, income, age, marital status, and occupation. Independent of the specific election or country examined, these results have been confirmed time and time again by subsequent studies. (@smets2013embarrassment) Smets and Van Ham (2012) conducted a thorough meta-analysis on individual level turnout and also found that education, age, income, and marital status were consistently statistically significant in regards to voter turnout. It should, however, be noted that there are substantial limits in the predictive and explanatory power of even well-specified statistical models. As Matsusaka (1999) has shown, voter turnout models generally suffer from a low r-squared, meaning they explain only a limited percentage of the factors behind the phenomenon.(@matsusaka1999voter)    

It is important to note that other empirical works (Cox and Munger, 1989; Leighley and Nagler, 1992) have elaborated on these findings by focusing on systemic characteristics and have found that the closeness of the election, contextual socio-economic conditions,[11] and registration laws have an impact on who goes to the polls. In a comprehensive meta-analysis on aggregate-level studies[12], Geys (2006) found that population size and electoral closeness, population stability, and past turnout were consistently positively significant, whereas racial diversity was consistently negatively significant. The interplay between individual predictors and systemic predictors has yet to be fully explored.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Bibliography
